---
title: "Report on MovieLens Project"
author: "Anh N.H. Nguyen"
date: "10/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a report on the MovieLens project. The data was obtained from [the GroupLens research lab](https://grouplens.org/datasets/movielens/latest/).

We are going to use the following library:

```{r loading-libs, message=FALSE}
library(tidyverse)
library(caret)
```

## MovieLens 10M dataset and edex/validation partition

```{r dataset, echo=FALSE}
dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- read.table(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                      col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                           title = as.character(title),
                                           genres = as.character(genres))

movielens <- left_join(ratings, movies, by = "movieId")

set.seed(1, sample.kind = "Rounding")
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")

removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)
```

## Overview

edX dataset includes 6 features of 9,000,055 observations. Validation dataset has similar features of 999,999 observations. 

```{r overview, echo=FALSE}
glimpse(edx)
glimpse(validation)
summary(edx)
```

## Ratings General

No user gives 0 as a rating. Half-star ratings are less common than whole-star ratings. Top 5 ratings are: 4, 3, 5, 3.5 and 2. 

```{r ratings-general, echo=FALSE}
edx %>%
  group_by(rating) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = rating, y = count)) +
  geom_line()
```

## Ratings by Genres

Top 5 rated genres are: Drama, Comedy, Action, Thriller, and Adventure

```{r ratings-genres, echo=FALSE}
edx %>% separate_rows(genres, sep = "\\|") %>%
  group_by(genres) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

## Movies Rankings

Top rated movies are: Pulp Fiction, Forrest Gump, Silence of the Lambs, Jurassic Park, etc. 

```{r movies, echo=FALSE}
edx %>% group_by(movieId, title) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```